---
title: "Single Cell Xenopus TMT Analysis2"
author: "Kyle Kovary"
date: "5/15/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RColorBrewer)
library(pheatmap)
library(GGally)
library(VIM)
library(org.Xl.eg.db)
library(GO.db)
library(ggrepel)
library(plotly)

# Import Data
files = list.files('TMTdata/', pattern = 'TargetPeptideSpectrumMatch', full.names = T)

for(i in seq_along(files)){
  if(i == 1 & exists('pepRaw')){
    rm(pepRaw)
  } else if(!exists('pepRaw')){
    pepRaw = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
  } else{
    temp = read_tsv(file = files[i]) %>% mutate(file = files[i], massID = 1:nrow(.))
    colnames(temp) = colnames(pepRaw)
    pepRaw = rbind(pepRaw, temp)
    rm(temp)
  }
}
pepRaw <- pepRaw %>% dplyr::select(file, massID, everything()) %>% filter(!is.na(`126`), 
                                                                  `Quan Usage` == 'Use',
                                                                  Confidence != 'Low')

```

## Data Organization


```{r }
pepRaw = pepRaw %>% 
  mutate(run = c('ctrl','a','b','c','d','e')[match(pepRaw$file, files)]) %>% 
  group_by(file, `Annotated Sequence`) %>% 
  mutate(pepID = letters[1:n()]) %>% ungroup() %>%
  dplyr::select(file, run, pepID, everything())

pepTidy = pepRaw %>% 
  gather(key = label, value = intensity, '126':'131') %>% 
  dplyr::select(run, label, intensity, everything(), -file) %>% 
  mutate(condition = ifelse(run == 'ctrl', 'ctrl',c('ctrl',0,20,40,60,80)[match(label,c(126:131))]),
         intensity = as.numeric(intensity)) %>% unite(peptide, c(`Annotated Sequence`,'pepID')) %>%
  dplyr::rename(Uniprot = `Master Protein Accessions`) %>% 
  unite(cell, c('run','label','condition'), remove = F) %>%
  dplyr::select(cell, run, label, condition, peptide, Uniprot, intensity)
```

## Variance and Co-variance Between Cells Before Normalization


```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
# ggplot(pepTidy, aes(log2(intensity), colour = cell)) + 
#   geom_density() + theme_bw() + xlim(-20,20)
```

The peptide intensities appear to follow a log-normal distribution.

```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
pepTidy %>% filter(condition == 'ctrl') %>% 
  dplyr::select(cell, peptide, intensity) %>% spread(cell, intensity) %>%

# ggpairs(log2(pairs[,2:ncol(pairs)]), 
#         lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
```

```{r echo=TRUE, fig.height=4, fig.width=4,message=FALSE, warning=FALSE}
pepTidy %>% filter(condition != 'ctrl') %>% 
  mutate(intensity = log2(intensity)) %>%
  group_by(peptide) %>% 
  summarise(mean = mean(intensity, na.rm = T), 
            median = median(intensity, na.rm = T), 
            sd = sd(intensity, na.rm = T),
            cv = sd / mean) %>%
ggplot(., aes(x = median, y = cv)) + 
  geom_point(size = 0.25, alpha = 0.1) + 
  geom_smooth(method = lm, formula = y ~ poly(x, 2)) +
  ylim(0,0.2)
```

## Variance and Co-variance Between Cells After Removing Peptides Outside of 2*sd

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
pepTidy <- pepTidy %>% dplyr::filter(abs(log2(intensity)) < 2*sd(log2(pepTidy$intensity), na.rm = T))

#ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-4,4)
```

The standard deviation of all peptides across eggs was calculated, and every peptide that had a value outside of `r 2*sd(log2(pepTidy$intensity), na.rm = T)`

## Group into proteins based on median

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
protTidy = pepTidy %>% group_by(cell, run, label, condition, Uniprot) %>% 
  summarise(abundance = median(intensity, na.rm = T)) %>% ungroup()

pairs = protTidy %>% dplyr::filter(condition == 'ctrl') %>% 
  dplyr::select(cell, Uniprot, abundance) %>% 
  spread(cell, abundance)

ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()

pairs = protTidy %>% dplyr::select(cell, Uniprot, abundance) %>% spread(cell, abundance)

```

## Impute data
```{r}
protImpute = kNN(data = pairs, k = 1) %>% as_tibble() %>% dplyr::select(Uniprot:e_131_80)

protVar <- protImpute %>% gather(key = 'cell', value = 'abundance', 2:ncol(protImpute)) %>%
  separate(cell, into = c('run','label','condition'), remove = F) %>% 
  filter(condition != 'ctrl') %>% group_by(Uniprot) %>%
  summarise(var = var(abundance),
            sd = sd(abundance),
            mean = mean(abundance),
            cv = sd / mean)

ggplot(protVar, aes(x = sd)) + geom_histogram(bins = 50) + theme_bw()
```

## Variance vs Mean
```{r}
ggplot(protVar, aes(x = log2(mean), y = var/mean^2)) + geom_point() + geom_smooth(method = lm, formula = y ~ poly(x, 2))
```


## Calculate variance and covariance
```{r}
cor = cor(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
colnames(cor) = protImpute$Uniprot
rownames(cor) = protImpute$Uniprot

cov = cov(t(log2(protImpute[c(1:ncol(protImpute))[-c(1,grep('ctrl',colnames(protImpute)))]])), use = 'pairwise.complete.obs')
colnames(cov) = protImpute$Uniprot
rownames(cov) = protImpute$Uniprot


mypal = rev(colorRampPalette(brewer.pal(11, 'RdBu'))(201))
k = 2
# map = pheatmap(cor, show_colnames = F, show_rownames = F, color = mypal, breaks = seq(-1,1, by = 0.01), cutree_cols = k, cutree_rows = k, legend = F)
#map

tibble(Uniprot = protImpute$Uniprot, Cluster = cutree(map$tree_row, k = k)) %>% write_csv('/Users/kylekovary/Downloads/Map.csv')
```

## PCA Analysis
```{r}
remove = which(is.na(pairs), arr.ind = T)[,1] %>% unique()
pca = prcomp(protImpute[,2:ncol(protImpute)])

pcaDF = tibble(cell = rownames(pca$rotation), PC1 = pca$rotation[,1], PC2 = pca$rotation[,2])
pcaDF = pcaDF %>% separate(cell, into = c('run','label','condition'), remove = F) %>% print()
ggplot(pcaDF, aes(x = PC1, y = PC2, colour = condition)) + geom_point(size = 5)
```


## Variance and Co-variance Between Cells After Cell Normalization

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
pepTidy = pepTidy %>% group_by(cell) %>% mutate(intensity = intensity / median(intensity, na.rm = T))

# ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-8,8)
```

```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
pairs = pepTidy %>% filter(condition == 'ctrl') %>% 
  dplyr::select(cell, peptide, intensity) %>% spread(cell, intensity)

# ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
```


## Variance and Co-variance Between Cells After Peptide Batch Normalization

```{r echo=TRUE, fig.height=5, fig.width=8, message=FALSE, warning=FALSE}
pepTidy = pepTidy %>% group_by(run, peptide) %>% 
  mutate(intensity = intensity / median(intensity, na.rm = T)) %>% ungroup()
 
# ggplot(pepTidy, aes(log2(intensity), colour = cell)) + geom_density() + theme_bw() + xlim(-8,8)
```

```{r echo=TRUE, fig.height=10, fig.width=10,message=FALSE, warning=FALSE}
pairs = pepTidy %>% dplyr::filter(condition == 'ctrl') %>% 
  dplyr::select(cell, peptide, intensity) %>% spread(cell, intensity)
 
# ggpairs(log2(pairs[,2:ncol(pairs)]), lower = list(continuous = wrap("points", alpha = 0.1, size = 0.1))) + theme_bw()
```
